\section{Conclusion \& References}
\label{conclusion}

The development of FeedApp successfully integrated key technologies to create a functional and scalable application that efficiently manages polls and votes. Among the chosen technologies, \textbf{Java Servlets} stood out as the featured technology, providing detailed control over the system's logic and seamless handling of user requests. Servlets enabled the backend to efficiently process requests, interact with the persistence layer, and manage data flow across the application.

To complement Servlets, \textbf{RabbitMQ} played a supporting role in enabling asynchronous, real-time messaging for analytical data synchronization, while \textbf{SQLite} and \textbf{MongoDB} were utilized for data storage. This hybrid approach allowed SQLite to manage operational data efficiently and MongoDB to handle analytical data with flexibility and scalability.

For future improvements, adopting frameworks like \textbf{Spring Boot} could simplify backend management for more complex projects while enhancing the authentication system to improve security. Additionally, implementing advanced features such as interactive data visualizations for analytics could significantly boost the system's usability and provide more value to users.

Overall, this project established a robust foundation for further development while providing the team with valuable experience in utilizing modern technologies.

\subsection*{References}

\begin{itemize}
    \item Oracle Documentation: \textbf{``Java Servlet Technology''}: \url{https://docs.oracle.com/javaee/7/tutorial/servlets.htm}
    \item Mozilla Developer Network (MDN): \textbf{``CSS: Cascading Style Sheets''}: \url{https://developer.mozilla.org/en-US/docs/Web/CSS}
    \item Mozilla Developer Network (MDN): \textbf{``JavaScript Guide''}: \url{https://developer.mozilla.org/en-US/docs/Web/JavaScript/Guide}
    \item DB Browser for SQLite: \textbf{``Database Management Tool for SQLite''}: \url{https://sqlitebrowser.org/}
    \item MongoDB Documentation: \textbf{``Getting Started with MongoDB''}: \url{https://www.mongodb.com/docs/manual/}
    \item RabbitMQ Documentation: \textbf{``Messaging with RabbitMQ''}: \url{https://www.rabbitmq.com/documentation.html}
    \item Spring Boot Documentation: \textbf{``Spring Boot Reference Guide''}: \url{https://docs.spring.io/spring-boot/docs/current/reference/html/}
\end{itemize}

\subsection*{Github Link of the Project Code:}
\url{https://github.com/r4f4777/FeedApp/tree/main/Project%20Code }

\subsection*{Github Link of the SQL database:}
\url{https://github.com/r4f4777/FeedApp/tree/main/Databases }